Racial and Ethnic Differences in ETI

### 2024-03-18 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
ETI survival analysis

---
title: "Racial and Ethnic Differences in ETI"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    number-sections: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
bibliography: /Users/timvigers/Documents/Miscellaneous/zotero.bib
csl: /Users/timvigers/Documents/Miscellaneous/american-medical-association.csl
editor: source
---

```{r setup}
#| include: false
# Libraries
library(segmented)
library(Hmisc)
library(tidyverse)
library(gtsummary)
library(lmerTest)
library(nlme)
library(broom.mixed)
library(emmeans)
library(gt)
library(performance)
library(mcp)
library(ggsurvfit)
# Paths
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

```{r results='asis'}
invisible(lapply(c("2019", "2020", "2021"), function(n) {
  cat("\n")
  cat("###", n)
  cat("\n")
  cat("\n")
  cat("#### All patients")
  cat("\n")
  d <- years[[n]]
  t <- d %>%
    select(eti_elig, contains("Vx"), race) %>%
    tbl_summary(
      by = race, missing_text = "Missing",
      type = all_dichotomous() ~ "categorical"
    ) %>%
    add_p(test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)) %>%
    separate_p_footnotes() %>%
    bold_labels()
  print(t)
  cat("\n")
  cat("\n")
  cat("#### Eligible and prescribed off-label only")
  cat("\n")
  t <- d %>%
    filter(eti_elig == "Yes" | Vx445comb == "Yes") %>%
    select(eti_elig, contains("Vx"), race) %>%
    tbl_summary(
      by = race, missing_text = "Missing",
      type = all_dichotomous() ~ "categorical"
    ) %>%
    add_ci() %>%
    add_p(test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)) %>%
    separate_p_footnotes() %>%
    bold_labels()
  print(t)
  cat("\n")
}))
```

@@

```{r}
# Create survival dataset
surv_dat <- encounter %>%
  mutate(stop = as.numeric(encounterdate - eti_elig_date)) %>%
  filter(stop >= 0) %>%
  group_by(eDWID) %>%
  mutate(start = lag(stop, default = 0)) %>%
  select(eDWID, race, encounterdate, encounterage, eti_elig_date, Vx445comb, start, stop) %>%
  filter(stop <= first(stop[Vx445comb == "Yes"]))
surv_dat$Vx445comb <- as.numeric(factor(surv_dat$Vx445comb, levels = c("No", "Yes"))) - 1
# Fit
fit <- survfit(Surv(start, stop, Vx445comb, type = "interval") ~ race,
  id = eDWID, data = surv_dat
)
# Plot
names(fit$strata) <- gsub("race=", "", names(fit$strata))
fit %>%
  ggsurvfit(type = "risk") +
  add_confidence_interval() +
  add_risktable(
    risktable_stats =
      c("{round(n.risk)}", "{round(cum.event)}"),
    stats_label = c("Eligible", "Num. Started")
  ) +
  add_risktable_strata_symbol() +
  add_quantile() +
  ylab("Proportion on Vx445comb") + xlab("Days From Eligibility")
```

@@

```{r}
fit <- survfit(Surv(start, stop, Vx445comb, type = "interval") ~ race,
  id = eDWID, data = surv_dat[surv_dat$race != "Asian", ]
)
# Plot
names(fit$strata) <- gsub("race=", "", names(fit$strata))
fit %>%
  ggsurvfit(type = "risk") +
  add_confidence_interval() +
  add_risktable(
    risktable_stats =
      c("{round(n.risk)}", "{round(cum.event)}"),
    stats_label = c("Eligible", "Num. Started")
  ) +
  add_risktable_strata_symbol() +
  add_quantile() +
  ylab("Proportion on Vx445comb") + xlab("Days From Eligibility")
```

>>>>

---
title: "Racial and Ethnic Differences in ETI"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    number-sections: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
editor: source
---

```{r setup}
#| include: false
# Libraries
library(segmented)
library(Hmisc)
library(tidyverse)
library(gtsummary)
library(lmerTest)
library(nlme)
library(broom.mixed)
library(emmeans)
library(gt)
library(performance)
library(mcp)
library(survival)
library(ggsurvfit)
library(ggstats)
library(naniar)
# Paths
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

```{r results='asis'}
invisible(lapply(c("2019", "2020", "2021", "2022"), function(n) {
  cat("\n")
  cat("###", n)
  cat("\n")
  cat("\n")
  cat("#### All patients")
  cat("\n")
  d <- years[[n]]
  t <- d %>%
    select(eti_elig, contains("Vx"), race) %>%
    tbl_summary(
      by = race, missing_text = "Missing",
      type = all_dichotomous() ~ "categorical"
    ) %>%
    add_p(test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)) %>%
    separate_p_footnotes() %>%
    bold_labels()
  print(t)
  cat("\n")
  cat("\n")
  cat("#### Eligible and prescribed off-label only")
  cat("\n")
  t <- d %>%
    filter(eti_elig == "Yes" | Vx445comb == "Yes") %>%
    select(eti_elig, contains("Vx"), race) %>%
    tbl_summary(
      by = race, missing_text = "Missing",
      type = all_dichotomous() ~ "categorical"
    ) %>%
    add_ci() %>%
    add_p(test.args = all_tests("fisher.test") ~ list(simulate.p.value = T)) %>%
    separate_p_footnotes() %>%
    bold_labels()
  print(t)
  cat("\n")
}))
```

@@

```{r}
# Create survival dataset
surv_dat <- encounter %>%
  mutate(stop = as.numeric(encounterdate - eti_elig_date)) %>%
  filter(stop > 0) %>%
  group_by(eDWID) %>%
  mutate(start = lag(stop, default = 0)) %>%
  select(
    eDWID, race, Insurance, family_income, patient_education, father_education,
    mother_education, encounterdate, encounterage, eti_elig_date, Vx445comb,
    start, stop
  ) %>%
  filter(stop <= first(stop[Vx445comb == "Yes"]))
surv_dat$Vx445comb <- as.numeric(factor(surv_dat$Vx445comb, levels = c("No", "Yes"))) - 1
# Fit
y <- "Surv(time = start, time2 = stop, event = Vx445comb)"
surv_f <- as.formula(paste0(y, "~race"))
fit <- survfit(surv_f, id = eDWID, data = surv_dat)
```

### Days to percentile started on ETI

```{r}
# Quantiles
tbl_survfit(fit, probs = c(0.25, 0.5, 0.75))
```

### Plots

```{r}
# Plot
names(fit$strata) <- gsub("race=", "", names(fit$strata))
fit %>%
  ggsurvfit(type = "risk") +
  add_confidence_interval() +
  add_risktable(
    risktable_stats =
      c("{round(n.risk)}", "{round(cum.event)}")
  ) +
  add_risktable_strata_symbol() +
  add_quantile() +
  ylab("Proportion on Vx445comb") + xlab("Days From Eligibility")
```

@@

```{r}
fit <- survfit(surv_f, id = eDWID, data = surv_dat[surv_dat$race != "Asian", ])
# Plot
names(fit$strata) <- gsub("race=", "", names(fit$strata))
fit %>%
  ggsurvfit(type = "risk") +
  add_confidence_interval() +
  add_risktable(
    risktable_stats =
      c("{round(n.risk)}", "{round(cum.event)}"),
    stats_label = c("Eligible", "Num. Started")
  ) +
  add_risktable_strata_symbol() +
  add_quantile() +
  ylab("Proportion on Vx445comb") + xlab("Days From Eligibility")
```

### Cox PH models

#### Base model (unadjusted)

```{r}
cox_res <- coxph(surv_f, data = surv_dat, cluster = eDWID)
cox_res %>%
  tbl_regression() %>%
  add_n(location = "level") %>%
  add_nevent(location = "level")
ggcoef_model(cox_res,
  show_p_values = F, add_reference_rows = F,
  signif_stars = F, colour = F,
)
```

#### Adjusted for insurance status, family income, and education

```{r}
new_surv_f <- update(surv_f, . ~ . + Insurance + family_income +
  patient_education + father_education + mother_education)
cox_res <- coxph(new_surv_f, data = surv_dat, cluster = eDWID)
cox_res %>%
  tbl_regression() %>%
  add_n(location = "level") %>%
  add_nevent(location = "level")
ggcoef_model(cox_res,
  show_p_values = F, add_reference_rows = F,
  signif_stars = F, colour = F,
)
```

##### Missing data

```{r}
gg_miss_var(surv_dat, show_pct = T, facet = race)
```

#### Adjusted for insurance status only

```{r}
new_surv_f <- update(surv_f, . ~ . + Insurance)
cox_res <- coxph(new_surv_f, data = surv_dat, cluster = eDWID)
cox_res %>%
  tbl_regression() %>%
  add_n(location = "level") %>%
  add_nevent(location = "level")
ggcoef_model(cox_res,
  show_p_values = F, add_reference_rows = F,
  signif_stars = F, colour = F,
)
```

#### Adjusted for combined insurance status only

Health Insurance Policy (e.g. Private Insurance) = Private

Indian Health Service = Public

Medicaid = Public

Medicare = Public

More than one = More than one

No Insurance = No Insurance

Other = Other

State special needs program, e.g., BCMH, CCS, CRS, GHPP, etc. = Other

TriCare or other military health plan = Tricare

```{r}
surv_dat$Insurance <- factor(surv_dat$Insurance,
  levels = c(
    "Health Insurance Policy (e.g. Private Insurance)",
    "Indian Health Service",
    "Medicaid",
    "Medicare",
    "More than one",
    "No Insurance",
    "Other",
    "State special needs program, e.g., BCMH, CCS, CRS, GHPP, etc.",
    "TriCare or other military health plan"
  ),
  labels = c(
    "Private", "Public", "Public", "Public", "More Than One", "No Insurance",
    "Other", "Other", "Tricare"
  )
)
cox_res <- coxph(new_surv_f, data = surv_dat, cluster = eDWID)
cox_res %>%
  tbl_regression() %>%
  add_n(location = "level") %>%
  add_nevent(location = "level")
ggcoef_model(cox_res,
  show_p_values = F, add_reference_rows = F,
  signif_stars = F, colour = F,
)
```
